Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-02-12 ms.tex
Update Fisher matrix diagonalization section

\documentclass{iopart}
\usepackage{iopams}
\usepackage{bib/aas_macros}
\usepackage[colorlinks]{hyperref}
\input{ligo-acronyms/acronyms}

@@

@@

@@

The Fisher information matrix for a measurement $\tilde x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
%
\begin{equation}\label{eq:general-fisher-matrix}
	\mathcal{I}_{jk} = \mathrm{E} \, \left[
		-\frac{\partial^2 \log
			\mathcal{L}(\tilde x ; \boldsymbol\theta)}
			{\partial \theta_j \theta_k}
	\right].
\end{equation}
%
This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
%
\begin{equation}\label{eq:gaussian-fisher-matrix}
	\mathcal{I}_{jk} = \int_0^\infty \Re \left[
        \left(\frac{\partial \tilde h}{\partial \theta_j}\right)^*
        \left(\frac{\partial \tilde h}{\partial \theta_k}\right)
	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
\end{equation}
%
This form is more useful because it involves only first derivatives, and of the signal $\tilde h (\omega)$ rather than the entire observation $\tilde x (\omega)$. In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
%
\begin{equation}\label{eq:angular-frequency-moments}
    \overline{\omega^k} =
        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
\end{equation}
%
the Fisher matrix for the signal model in \Eref{eq:signal-model} is
%
\begin{eqnarray*}
	\mathcal{I} = \bordermatrix{
        ~ & \rho & \gamma & \tau \cr
        \rho & 1 & 0 & 0 \cr
        \gamma & 0 & \rho^2 & -\rho^2 \overline{\omega} \cr
        \tau & 0 & -\rho^2 \overline{\omega} & \rho^2 \overline{\omega^2}
    }.
\end{eqnarray*}

\section{Posterior distribution}

@@

\marginpar{This part is correct, but uses the old notation for the frequency moments. Convert to the new notation. Also, add physical interpretation to each change of variables.}%
%
Now we will perform four changes of variables. The first change of variables is from $C$ to $\rho = \sqrt{M_0}C$. This change is intended to make the variance of the absolute value of the amplitude equal to one. 
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & \rho^2 & -2\pi\rho^2 N_1 \\
 0 & -2\pi\rho^2 N_1 & 4\pi^2 \rho^2 N_2
 \end{array} } \right)
\]

The second from $t$ to $\phi_t = t2\pi \sqrt{N_2}$ is meant to get rid of $t_0$,
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & \rho^2 & -\rho^2\frac{N_1}{\sqrt{N_2}} \\
 0 & -\rho^2\frac{N_1}{\sqrt{N_2}} & \rho^2
 \end{array} } \right)
\]

The third introduces two new variables, $\phi_s = \frac{1}{\sqrt{2}}(\phi + \phi_t)$ and $\phi_d = \frac{1}{\sqrt{2}}(\phi - \phi_t)$. This change of variables diagonalizes the matrix.
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & (1 - \frac{N_1}{\sqrt{N_2}})\rho^2 & 0 \\
 0 & 0 & (1 + \frac{N_1}{\sqrt{N_2}})\rho^2
 \end{array} } \right)
\]

Now we'll do a final change of variables to make the phase parameters have the same variance. We let $\alpha_s = \phi_s \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ and $\alpha_d = \phi_d \sqrt{1 - \frac{N_1}{\sqrt{N_2}}}$ which gives us the matrix
\[
\mathcal{I} =
\left( {\begin{array}{ccc}
 1 & 0 & 0 \\
 0 & \rho^2 & 0 \\
 0 & 0 & \rho^2
 \end{array} } \right)
\]

>>>>

\documentclass{iopart}
\usepackage{amsopn}
\usepackage{iopams}
\usepackage{bib/aas_macros}
\usepackage[colorlinks]{hyperref}
\input{ligo-acronyms/acronyms}

@@

\DeclareMathOperator{\cov}{cov}

@@

\subsection{Fisher matrix}

@@

The Fisher information matrix for a measurement $\tilde x$ described by the unknown parameter vector $\boldsymbol{\theta}$ is
%
\begin{equation}\label{eq:general-fisher-matrix}
	\mathcal{I}_{jk} = \mathrm{E} \, \left[
		-\frac{\partial^2 \log
			\mathcal{L}(\tilde x ; \boldsymbol\theta)}
			{\partial \theta_j \theta_k}
	\right].
\end{equation}
%
This equation involves double derivatives, but substantial simplification is possible when---as in this case---the likelihood is Gaussian:
%
\begin{equation}\label{eq:gaussian-fisher-matrix}
	\mathcal{I}_{jk} = \int_0^\infty \Re \left[
        \left(\frac{\partial \tilde h}{\partial \theta_j}\right)^*
        \left(\frac{\partial \tilde h}{\partial \theta_k}\right)
	\right] \frac{1}{S(\omega)} \, \mathrm{d}\omega.
\end{equation}
%
This form is more useful because it involves only first derivatives, and of the signal $\tilde h (\omega)$ rather than the entire observation $\tilde x (\omega)$. In terms of the $k$th \ac{SNR}-weighted moment of angular frequency,
%
\begin{equation}\label{eq:angular-frequency-moments}
    \overline{\omega^k} =
        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \omega^k \, \mathrm{d}\omega \right]
        \left[ \int_0^\infty \frac{|\tilde u (\omega)|^2}{S(\omega)} \, \mathrm{d}\omega \right]^{-1},
\end{equation}
%
the Fisher matrix for the signal model in \Eref{eq:signal-model} is
%
\begin{equation}\label{eq:fisher-matrix}
	\mathcal{I} = \bordermatrix{
        ~ & \rho & \gamma & \tau \cr
        \rho & 1 & 0 & 0 \cr
        \gamma & 0 & \rho^2 & -\rho^2 \overline{\omega} \cr
        \tau & 0 & -\rho^2 \overline{\omega} & \rho^2 \overline{\omega^2}
    }.
\end{equation}

\subsection{\acl{CRLB} and physical interpretation}

The \ac{CRLB} says that, element by element,
%
\begin{equation}
    \cov{
        \left(
        \begin{array}{c}
            \hat{\rho} - \rho \\
            \hat{\gamma} - \gamma \\
            \hat{\tau} - \tau
        \end{array}
        \right)
    } \geq \mathcal{I}^{-1}
\end{equation}
%
where a hat $\hat{\phantom{\rho}}$ denotes the estimate of a parameter. The Fisher matrix in \Eref{eq:fisher-matrix} is block diagonal, which implies that estimation errors in the signal amplitude $\rho$ are uncorrelated with the phase $\gamma$ and time $\tau$. A sequence of two changes of variables lends some physical interpretation to the nature of the coupled estimation errors in $\gamma$ and $\tau$.

First, we put the phase and time on the same footing by measuring the time in units of $1 / 2 \pi \sqrt{\overline{\omega^2}}$ with a change of variables from $\tau$ to $\gamma_\tau = 2 \pi \sqrt{\overline{\omega^2}} \tau$:
%
\begin{equation}
    \mathcal{I}' = \bordermatrix{
        ~ & \rho & \gamma & \gamma_\tau \cr
        \rho & 1 & 0 & 0 \cr
        \gamma & 0 & \rho^2 & -\rho^2\frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}} \cr
        \gamma_\tau & 0 & -\rho^2\frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}} & \rho^2
    }.
\end{equation}

The second change of variables, from $\gamma$ and $gamma_\tau$ to $\gamma_\pm = \frac{1}{\sqrt{2}}(\gamma \pm \gamma_\tau)$, diagonalizes the Fisher matrix:
%
\begin{equation}
    \mathcal{I}'' = \bordermatrix{
        ~ & \rho & \gamma_+ & \gamma_- \cr
        \rho & 1 & 0 & 0 \cr
        \gamma_+ & 0 & \left(1 - \frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}}\right)\rho^2 & 0 \cr
        \gamma_- & 0 & 0 & \left(1 + \frac{\overline{\omega}}{\sqrt{\overline{\omega^2}}}\right)\rho^2
    }.
\end{equation}
%
Thus, in the appopriate time units, the \textit{sum and difference} of the phase and time of the signal are measured independently.

@@
